Impact of Satellite Constellations on the Legacy Survey of Space and Time

### 2022-10-12 ms.tex
Updates from Overleaf

\author[0000-0002-8400-1910]{Jinghan Alina Hu}
\affiliation{Harvey Mudd College, Claremont, CA, USA}
\author[0000-0003-1305-7308]{Meredith L. Rawls}
\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
\author[0000-0003-2874-6464]{Peter Yoachim}
\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
\author{Zeljko Ivezic}
\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
\author{Kresimir Mihic}

@@

\section{Methodology}
We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and StarlinkV2 (29,988 satellites, altitude 340-614 km). All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more. These simulated satellite constellations are illustrated in Figure \ref{fig-simulated-constellations}.

@@

XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year. As a starting point we use the baseline scheduler simulation in \citet{yoachim2022}.

\begin{figure}[ht!]
\epsscale{0.37}
\plotone{plots/starlink_tles_v1_0.00_basisfunc.pdf}
\plotone{plots/starlink_tles_v2_0.00_basisfunc.pdf}
\plotone{plots/oneweb_tles_0.00_basisfunc.pdf}
\epsscale{1}
\caption{Example of satellite avoidance maps constructed for the scheduler. The maps have been rotated so zenith is in the center of the image. By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images. \label{fig-simulated-scheduler}}
\end{figure}

As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward weighting than other constellations. Although the OneWeb constellation has fewer satellites than the StarlinkV2 constellation, the OneWeb constellation is at a higher altitude, meaning that its satellites will be illuminated for a longer portion of the night, and not just close to twilight. Therefore, OneWeb shows more negative weighting than StarlinkV2 though it has a smaller satellite population. To investigate whether the scheduler behaves how we expect with the new satellite weighting added we create a testing function that measures the length of satellite streaks in the simulated field of views. To ensure efficiency, only satellites that are above the altitude limit and illuminated by the sun will be considered. Satellites below the altitude limit (indicated in the gray region in Figure \ref{fig-simulated-scheduler}) can’t be observed in pointings and are therefore not included for efficiency. For each satellite, we first determine whether they are in the field of view for any given telescope pointing by calculating their distance from the center of the field of view. If their distance from the field of view is less than the radius of the field of view, the satellite has crossed through the pointing. To quantify the impact of the satellite on the pointing, we then project the satellite location as well as the pointing to a 2D x,y plane. In the 2D plane, the field of pointing becomes roughly a circle and the start and end locations of the satellite become two points on the plane. Using the shapely python library, we calculate the intersection length of the satellite location points and the circle. Therefore, with a given simulated satellite constellation and a schedule of observations, we record the number of satellites that occur in each field of view and measure how long the satellite streaks are. Using a realistic estimated satellite streak width of 50 pixels ($\sim10$ arcseconds), we can convert the streak length results into the number of affected image pixels, which allows us to quantitatively measure the efficiency of the dodging algorithm with pixel loss.

\section{Results}
We begin by investigating the efficiency of the added weighting algorithm on dodging satellites in observations. As shown in Figure \ref{fig-pixel-loss-weight}, using a conservative estimate of 50 pixels as streak width, we found that higher dodging weights reduces pixels lost to satellite streaks. We also found that smaller constellations at lower orbital altitudes (StarlinkV1 for example) inherently results in less pixel loss per pointing, nearly independent of the dodging weight. Figure \ref{fig-pixel-loss-weight} shows that the added weighted dodging algorithm was able to effectively avoid satellite streaks in pointings.

@@

We then investigate the relationship between number of exposures taken and the dodging weight. As shown in Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer exposures, most likely due to longer slew times. With the additional avoidance weighting, the telescope may be prompted to slew to a location other than the desired pointing location and then slew back in order to avoid some satellites, resulting in fewer overall exposures. We also found that larger constellations like StarlinkV2 tend to decrease the number of exposures slightly more than smaller constellations (StarlinkV1), which is expected. More satellites or satellites at higher orbital altitudes result in bigger areas of avoidance on the sky, which results in more slewing to avoid the affected areas, which subsequently reduces exposures. \\

>>>>

\author[0000-0002-8400-1910]{Jinghan Alina Hu}
\affiliation{Harvey Mudd College, Claremont, CA, USA}
\author[0000-0003-1305-7308]{Meredith L. Rawls}
\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
\author[0000-0003-2874-6464]{Peter Yoachim}
\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
\author{Zeljko Ivezic}
\affiliation{Department of Astronomy and DiRAC, University of Washington, Seattle, WA, USA}
%% TODO: Add Kreso?

@@

\section{Methodology}
We begin by creating realistic forecasts of three different commercial satellite constellations: StarlinkV1 (4,408 satellites, altitude 540-570 km), OneWeb (6,372 satellites, altitude 1200 km), and Starlink Gen2 (29,988 satellites, altitude 340-614 km). All three of these constellations are highly likely to launch as designed (XXX-lol, no), or close to it, as both operators presently have functional satellites in orbit and are working to build and launch more. These simulated satellite constellations are illustrated in Figure \ref{fig-simulated-constellations}.

@@

XXX--I'll add some text describing the baseline scheduler that we use--note that we only run for the first year. As a starting point we use the baseline scheduler simulation in \citet{yoachim2022b}.

\begin{figure}[ht!]
\epsscale{0.37}
\plotone{plots/starlink_tles_v1_0.00_basisfunc.pdf}
\plotone{plots/starlink_tles_v2_0.00_basisfunc.pdf}
\plotone{plots/oneweb_tles_0.00_basisfunc.pdf}
\epsscale{1}
\caption{Example of satellite avoidance maps constructed for the scheduler, at sun altitude -17.1 degrees. The maps have been rotated so zenith is in the center of the image. By varying the weight placed on these maps, the scheduler will more actively avoid regions of the sky where satellites could streak images. \label{fig-simulated-scheduler}}
\end{figure}

As shown in Figure \ref{fig-simulated-scheduler}, the predicted OneWeb constellation map has more negative reward weighting than other constellations. Although the OneWeb constellation has fewer satellites than the Starlink Gen2 constellation, the OneWeb constellation is at a higher altitude, meaning that its satellites will be illuminated for a longer portion of the night, and not just close to twilight. Therefore, OneWeb shows more negative weighting than Starlink Gen2 though it has a smaller satellite population. To investigate whether the scheduler behaves how we expect with the new satellite weighting added we create a testing function that measures the length of satellite streaks in the simulated field of views. To ensure efficiency, only satellites that are above the altitude limit and illuminated by the sun will be considered. Satellites below the altitude limit (indicated in the gray region in Figure \ref{fig-simulated-scheduler}) can’t be observed in pointings and are therefore not included for efficiency. For each satellite, we first determine whether they are in the field of view for any given telescope pointing by calculating their distance from the center of the field of view. If their distance from the field of view is less than the radius of the field of view, the satellite has crossed through the pointing. To quantify the impact of the satellite on the pointing, we then project the satellite location as well as the pointing to a 2D x,y plane. In the 2D plane, the field of pointing becomes roughly a circle and the start and end locations of the satellite become two points on the plane. Using the shapely python library, we calculate the intersection length of the satellite location points and the circle. Therefore, with a given simulated satellite constellation and a schedule of observations, we record the number of satellites that occur in each field of view and measure how long the satellite streaks are. Using a realistic estimated satellite streak width of 50 pixels ($\sim10$ arcseconds), we can convert the streak length results into the number of affected image pixels, which allows us to quantitatively measure the efficiency of the dodging algorithm with pixel loss.

\section{Results}
We begin by investigating the efficiency of the added weighting algorithm on dodging satellites in observations. As shown in Figure \ref{fig-pixel-loss-weight}, using a conservative estimate of 50 pixels as streak width, we found that higher dodging weights reduces pixels lost to satellite streaks. We also found that smaller constellations at lower orbital altitudes (Starlink Gen1 for example) inherently results in less pixel loss per pointing, nearly independent of the dodging weight. Figure \ref{fig-pixel-loss-weight} shows that the added weighted dodging algorithm was able to effectively avoid satellite streaks in pointings.

@@

We then investigate the relationship between number of exposures taken and the dodging weight. As shown in Figure \ref{fig-pixel-loss-weight}, higher dodging weight results in fewer exposures, most likely due to longer slew times. With the additional avoidance weighting, the telescope may be prompted to slew to a location other than the desired pointing location and then slew back in order to avoid some satellites, resulting in fewer overall exposures. We also found that larger constellations like Starlink Gen2 tend to decrease the number of exposures slightly more than smaller constellations (Starlink Gen1), which is expected. More satellites or satellites at higher orbital altitudes result in bigger areas of avoidance on the sky, which results in more slewing to avoid the affected areas, which subsequently reduces exposures. \\
